---
title: "Practice Model Selection"
author: "Alexandra Boville"
date: "2024-04-15"
output: html_document
---

Libraries Used in this Exercise
```{r}
library(tidyverse)
library(naniar)
library(AICcmodavg)
library(ggplot2)
library(jtools)
library(MuMIn)
```

Loading Dataset
```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/Mammal_lifehistories_v2.txt"

d <- read_tsv(f, col_names = TRUE)

d
```


# Step 1: Replacing -999 values with NA
```{r}
d %>% replace_with_na_all(condition = ~.x == -999)
```
# Steps 2 + 3: Droping the variables litter size and refs and log transforming all remaining variables

```{r}
d <- d %>% 
  subset(select = -c(refs, `litter size`)) %>% 
  mutate(
    logMass = log(`mass(g)`),
    logGestation = log(`gestation(mo)`), 
    logNewborn = log(`newborn(g)`), 
    logWeaning = log(`weaning(mo)`),
    logWean.Mass = log(`wean mass(g)`),
    logAFR = log(`AFR(mo)`), 
    logMax.Life = log(`max. life(mo)`), 
    `logLitters/Year` = log(`litters/year`)
  )

d
```

# Step 4: Regressing Variables
```{r}
relGest <- resid(lm(logGestation ~ logMass, data = d, na.action = na.exclude))
relWean <- resid(lm(logWeaning ~ logMass, data = d, na.action = na.exclude))
relAFR <- resid(lm(logAFR ~ logMass, data = d, na.action = na.exclude))
relLife <- resid(lm(logMax.Life ~ logMass, data = d, na.action = na.exclude))
relNewbornMass <- resid(lm(logNewborn ~ logMass, data = d, na.action = na.exclude))
relWeaningMass <- resid(lm(logWean.Mass ~ logMass, data = d, na.action = na.exclude))

d$relGestation <- relGest
d$relWeaning <- relWean
d$relAFR <- relAFR
d$relLife <- relLife
d$relNewbornMass <- relNewbornMass
d$relWeaningMass <- relWeaningMass

d
```
# Step 5: Plotting Residuals

## Plotting residuals of max lifespan (relLife) in relation to Order
```{r}
ggplot(d, aes(x = order, y = relLife)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  theme_bw()
```

Primates, Rodentia and Xenarthra have the highest residuals of relative max lifespan. 

## Plotting residuals of newborn mass (relNewbornMass) in relation to Order

```{r}
ggplot(d, aes(x = order, y = relNewbornMass)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  theme_bw()
```

Cetacea and Rodentia have the highest residuals of relative newborn mass. 


## Plotting residuals of weaning mass (relWeaningMass) in relation to Order

```{r}
ggplot(d, aes(x = order, y = relWeaningMass)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  theme_bw()
```

Carnivora, Primates, Artiodactyla and Perissodactyla have the highest residuals for relative weaning mass. 

# Step 6: Running Models and Model Selection 

Droping NA values from dataset
```{r}
d <- d %>% 
  drop_na()

d
```

## Using Forward Selection to Determine which variables best predict max.life from predictor variables gestation(mo), newborn(g), weaning(mo), wean mass(g), litters/year, and overall body mass(g).

```{r}
m0 <- lm(data = d, logMax.Life ~ 1) #intercept only model
summary(m0)

add1(m0, scope = . ~ . + logGestation + logNewborn + logWeaning + logWean.Mass + `logLitters/Year` + logMass, test = "F")


m1 <- update(m0, formula = . ~ . + logGestation) #logGestation has the highest F value and the lowest RSS and the lowest AIC value so has the most explanatory power even though all predictor variables are significant 
summary(m1)

add1(m1, scope = . ~ . + logGestation + logNewborn + logWeaning + logWean.Mass + `logLitters/Year` + logMass, test = "F")

m2 <- update(m1, formula = . ~ . + `logLitters/Year`) #`logLitters/Year` has the highest F value and the lowest RSS and the lowest AIC value so has the most explanatory power
summary(m2)

add1(m2, scope = . ~ . + logGestation + logNewborn + logWeaning + logWean.Mass + `logLitters/Year` + logMass, test = "F")

m3 <- update(m2, formula = . ~ . + logMass) #logMass has the highest F value and the lowest RSS and the lowest AIC value so has the most explanatory power
summary(m3)

add1(m3, scope = . ~ . + logGestation + logNewborn + logWeaning + logWean.Mass + `logLitters/Year` + logMass, test = "F")

m4 <- update(m3, formula = . ~ . + logWeaning) #logWeaning has the highest F value and the lowest RSS and the lowest AIC value so has the most explanatory power
summary(m4)

add1(m4, scope = . ~ . + logGestation + logNewborn + logWeaning + logWean.Mass + `logLitters/Year` + logMass, test = "F")

m5 <- update(m4, formula = . ~ . + logNewborn) #logNewborn has the highest F value and the lowest RSS and the lowest AIC value so has the most explanatory power
summary(m5)

add1(m5, scope = . ~ . + logGestation + logNewborn + logWeaning + logWean.Mass + `logLitters/Year` + logMass, test = "F")

m6 <- update(m5, formula = . ~ . + logWean.Mass) #logWean.Mass has the highest F value and the lowest RSS and the lowest AIC value so has the most explanatory power
summary(m6)

```
## AICc values to determine best model
```{r}
aictab(list(m0, m1, m2, m3, m4, m5, m6), c("m0", "m1", "m2", "m3", "m4", "m5", "m6"))
```

Model 4 is selected to be the best! Models 4, 5 and 6 all have delta AICc values of less than 4. These top models include all variables but in relation to models 0-3 include logWeaning, logNewborn and logWean.Mass. 

## Using Forward Selection to Determine which variables best predict AFR from predictor variables gestation(mo), newborn(g), weaning(mo), wean mass(g), litters/year, and overall body mass(g).

```{r}
m2_0 <- lm(data = d, logAFR ~ 1) #intercept only model
summary(m2_0)

add1(m2_0, scope = . ~ . + logGestation + logNewborn + logWeaning + logWean.Mass + `logLitters/Year` + logMass, test = "F")


m2_1 <- update(m2_0, formula = . ~ . + `logLitters/Year`) #`logLitters/Year`  has the highest F value and the lowest RSS and the lowest AIC value so has the most explanatory power even though all predictor variables are significant 
summary(m2_1)

add1(m2_1, scope = . ~ . + logGestation + logNewborn + logWeaning + logWean.Mass + `logLitters/Year` + logMass, test = "F")

m2_2 <- update(m2_1, formula = . ~ . + logGestation) #logGestation has the highest F value and the lowest RSS and the lowest AIC value so has the most explanatory power
summary(m2_2)

add1(m2_2, scope = . ~ . + logGestation + logNewborn + logWeaning + logWean.Mass + `logLitters/Year` + logMass, test = "F")

m2_3 <- update(m2_2, formula = . ~ . + logWeaning) #logWeaning has the highest F value and the lowest RSS and the lowest AIC value so has the most explanatory power
summary(m2_3)

add1(m2_3, scope = . ~ . + logGestation + logNewborn + logWeaning + logWean.Mass + `logLitters/Year` + logMass, test = "F")

m2_4 <- update(m2_3, formula = . ~ . + logMass) #logMass has the highest F value and the lowest RSS and the lowest AIC value so has the most explanatory power
summary(m2_4)

add1(m2_4, scope = . ~ . + logGestation + logNewborn + logWeaning + logWean.Mass + `logLitters/Year` + logMass, test = "F")

m2_5 <- update(m2_4, formula = . ~ . + logWean.Mass) #logWean.Mass has the highest F value and the lowest RSS and the lowest AIC value so has the most explanatory power
summary(m2_5)

add1(m2_5, scope = . ~ . + logGestation + logNewborn + logWeaning + logWean.Mass + `logLitters/Year` + logMass, test = "F")

m2_6 <- update(m2_5, formula = . ~ . + logNewborn) #logNewborn has the highest F value and the lowest RSS and the lowest AIC value so has the most explanatory power
summary(m2_6)

```

## AICc values to determine best model
```{r}
aictab(list(m2_0, m2_1, m2_2, m2_3, m2_4, m2_5, m2_6), c("m2_0", "m2_1", "m2_2", "m2_3", "m2_4", "m2_5", "m2_6"))
```

Model 4 is again predicted to be the best model. Models 4, 3, 5 and 2 all have delta AICc values of less than 4. The variables that appear in these top models are logMass, logWeaning, logWean.Mass and logGestation. 

## Plotting Model Avg. Coefficients and CI's
```{r}
#Model-Average Coefficients and CIs for top models for max.life
m_maxlife <- lm(data = d, logMax.Life ~ logGestation + logNewborn + logWeaning + logWean.Mass + `logLitters/Year` + logMass, na.action = "na.fail")
ms_maxlife <- dredge(m_maxlife)
(mods.avg1 <- summary(model.avg(ms_maxlife, subset = delta < 4,fit = TRUE)))
confint
plot(mods.avg1, full = TRUE, intercept = FALSE)


#Model-Average Coefficients and CIs for top models for AFR
m_AFR <- lm(data = d, logAFR ~ logGestation + logNewborn + logWeaning + logWean.Mass + `logLitters/Year` + logMass, na.action = "na.fail")
ms_AFR <- dredge(m_AFR)
(mods.avg2 <- summary(model.avg(ms_AFR, subset = delta < 4,fit = TRUE)))
confint
plot(mods.avg2, full = TRUE, intercept = FALSE)



```



